HETDEX Paper

### 2015-06-04 ms.tex
worked on aardvark section

This support will allow Mr. Boada to contribute to the core science of my NSF grant (AST 1413317) for the DECam/IRAC Galaxy Environment Survey (DIRGES). He will studying both simulated and real data to understand the reliability in measuring galaxy cluster masses and dynamical states using spectroscopic information for emission-line galaxies. DIRGES will identify as many as 100 galaxy cluster progenitors at 0.8 $<$ z $<$ 1.5. Significant gains in the understanding of these clusters are possible if we can (1) interpret their dynamical nature through spectroscopy, and (2) if we can link them to clusters at lower redshift, z $<$ 0.5, from within the same dataset to suppress potential observational biases. I have plans to obtain the spectroscopy to achieve both of these supplemental goals, but the majority of the spectroscopic redshifts will be for emission-line galaxies. It is not straightforward to use these data to interpret the kinematic and dynamical states of the clusters. DIRGES is covered by SDSS Stripe 82 (Annis et al. 2014), the Dark Energy Survey (DES), and the Hobby Eberly Telescope Dark Energy eXperiment (HETDEX). One key is HETDEX, which is a spectroscopic survey covering $\sim$400 deg2 and measuring redshifts for as many as ∼106 emission-line galaxies at z $<$ 0.5. HETDEX will contain as many as 100 galaxy Virgo-sized clusters (Mdyn$\sim$1015 M⊙) out to z $<$ 0.5. And, because HETDEX covers DIRGES, the data will contain both low-redshift and high-redshift clusters in the same imaging dataset, allowing the most homogeneous study of the low-z and high-z universe yet.

\section{AARDVARK SIMULATION}
\editorial{Waiting to get a look at what Casey comes up with for the proposal we are supposedly writing. If he does a decent job then I'll try to expand it a bit and fill out this section.}
\begin{figure} 
	\includegraphics[width=0.49\textwidth]{surveyArea.pdf} 
	\caption{A polar azimuthal equidistant projection of the southern sky showing the sky area (approximately a quarter of the sky) covered by the Aardvark Simulations. Concentric circles are lines of constant declination, 10 to $-80\degree$ in $10\degree$ increments. Radial lines are lines of constant right ascension. Opaqueness of the region indicates the overlap of catalog tiles (see text for details).} 
	\label{fig:survey area} 
\end{figure}

\section{THE HETDEX SURVEY}\label{sec:Hetdex}
\editorial{This should probably go into the introduction. I don't think we need a whole section on something that a bunch of papers have already talked about. The text below is taken from Chiang 2015 and is NOT original to me.}

The Hobby Eberly Telescope Dark Energy Experiment (HETDEX; Hill et al. 2008b) is pioneering the instrumentation development and observations of high-redshift large scale structures using wide-field IFUs. In a 3-year baseline starting from late 2015, HETDEX will leverage the cosmic evolution of the dark energy equation of state with high redshift (z $>$ 2) constraints imprinted by the Baryon Acoustic Oscillations (BAO; Eisenstein 2005) in the early universe. The program will perform a redshift survey of LAEs in 300 deg2 (Spring field) plus 150 deg2 (Fall field) at 1.9 $<$ z $<$ 3.5 (with a filling factor of 1/4.5), with a total survey volume of $\sim$ 8.5 Gpc3. The survey uses the 10-m Hobby-Eberly Telescope (HET; Ramsey et al. 1998) with a wide-field upgrade to reach a 22x22 arcmin2 field-of-view. Blind spectroscopy (R $\sim$ 750 in 350–550 nm) with no pre-selection of targets will be performed using the Visible Integral-Field Replicable Unit Spectrograph (VIRUS; Hill et al. 2012, 2014). With the LAE redshifts, HETDEX will pinpoint numerous locations of the highest density concentrations at 1.9 $<$ z $<$ 3.5, generating a substantially large and homogeneous sample of cluster progenitors in the key epoch of cluster formation before virialization.

\section{Mock Observations}
\editorial{Here we need to talk about about how we set up the observations to look like what we think the HETDEX survey will see. It should be more technical than the general discussion about what the survey hopes to accomplish, that'll go in the intro. A zoomed figure showing the tiling footprints would be appropriate.}

@@

\subsection{Line of Sight Velocity Dispersion}
We first calculate the line of sight velocity (LOSV) to each galaxy, where
\begin{equation}
	LOSV = c\frac{z - z_c}{1+z_c}
\end{equation}
and $c$ is the speed of light in \kms, $z$ is the redshift of the individual galaxy, and $z_c$ is the overall cluster redshift described in the previous section.

@@

\subsection{Dynamical Mass}
Recently, the relationship between the LOSVD and dynamical mass has been the focus of several studies \citeeg{Evrard2008, Saro2013, Sifon2013, VanderBurg2014}, and a best fitting relationship for the mass enclosed by $r_{200c}$ of the form
\begin{equation}
	M_{200c} = \frac{10^{15}}{h(z)} \bigg{(}\frac{\sigma_{1D}}{A_{1D}} \bigg{)}^{-\alpha} \Msol
\end{equation}
with $A_{1D} =$ 1040-1140 \kms\ (\citealt{Munari2013}; referred to as $\sigma_{15}$ in \citealt{Evrard2008} and other works), $\alpha = 1/3$, $h(z) = H(z)/100$, and $\sigma_{1D}$ is the LOSVD of the velocity tracers (dark matter particles, subhalos or galaxies).

@@

\section{ {\rm[\ion{O}{ii}]} Luminosity}
HETDEX is designed to detect Lyman-$\alpha$ at redshift two. That allows us to detect [\ion{O}{ii}] ($\lambda\lambda 3727-3729\AAA$ doublet) emitters to $z\sim 0.49$. The catalog does not provide [\ion{O}{ii}] luminosities so we must assign them empirically. We use 503113 galaxies from the Sloan Digital Sky Survey (SDSS) Data Release 12 \citep{Alam2015} from $z = 0.05 - 0.2$, which are selected which have no redshift warning.

>>>>

This support will allow Mr. Boada to contribute to the core science of my NSF grant (AST 1413317) for the DECam/IRAC Galaxy Environment Survey (DIRGES). He will studying both simulated and real data to understand the reliability in measuring galaxy cluster masses and dynamical states using spectroscopic information for emission-line galaxies. DIRGES will identify as many as 100 galaxy cluster progenitors at 0.8 $<$ z $<$ 1.5. Significant gains in the understanding of these clusters are possible if we can (1) interpret their dynamical nature through spectroscopy, and (2) if we can link them to clusters at lower redshift, z $<$ 0.5, from within the same dataset to suppress potential observational biases. I have plans to obtain the spectroscopy to achieve both of these supplemental goals, but the majority of the spectroscopic redshifts will be for emission-line galaxies. It is not straightforward to use these data to interpret the kinematic and dynamical states of the clusters. DIRGES is covered by SDSS Stripe 82 \citep{Annis2014}, the Dark Energy Survey (DES), and the Hobby Eberly Telescope Dark Energy eXperiment (HETDEX). One key is HETDEX, which is a spectroscopic survey covering $\sim$400 deg$^2$ and measuring redshifts for as many as ∼106 emission-line galaxies at z $<$ 0.5. HETDEX will contain as many as 100 galaxy Virgo-sized clusters ($M_{dyn}\sim10^{15}\msol$) out to z $<$ 0.5. And, because HETDEX covers DIRGES, the data will contain both low-redshift and high-redshift clusters in the same imaging dataset, allowing the most homogeneous study of the low-z and high-z universe yet.

\editorial{This should probably go into the introduction. I don't think we need a whole section on something that a bunch of papers have already talked about. The text below is taken from Chiang 2015 and is NOT original to me.}

The Hobby Eberly Telescope Dark Energy Experiment (HETDEX; \citealt{Hill2008}) is pioneering the instrumentation development and observations of high-redshift large scale structures using wide-field IFUs. In a 3-year baseline starting from late 2015, HETDEX will leverage the cosmic evolution of the dark energy equation of state with high redshift (z $>$ 2) constraints imprinted by the Baryon Acoustic Oscillations (BAO; Eisenstein 2005) in the early universe. The program will perform a redshift survey of LAEs in 300 deg2 (Spring field) plus 150 deg2 (Fall field) at 1.9 $<$ z $<$ 3.5 (with a filling factor of 1/4.5), with a total survey volume of $\sim$ 8.5 Gpc3. The survey uses the 10-m Hobby-Eberly Telescope (HET; \citealt{Ramsey1998}) with a wide-field upgrade to reach a 22x22 arcmin2 field-of-view. Blind spectroscopy (R $\sim$ 750 in $350–550$ nm) with no pre-selection of targets will be performed using the Visible Integral-Field Replicable Unit Spectrograph (VIRUS; \citealt{Hill2012}). With the LAE redshifts, HETDEX will pinpoint numerous locations of the highest density concentrations at 1.9 $<$ z $<$ 3.5, generating a substantially large and homogeneous sample of cluster progenitors in the key epoch of cluster formation before virialization.

\section{Data and Observations}

Blah blah intro stuff...

\subsection{The ``Aardvark'' Catalogs}
\begin{figure} 
	\includegraphics[width=0.49\textwidth]{surveyArea.pdf} 
	\caption{A polar azimuthal equidistant projection of the southern sky showing the sky area (approximately a quarter of the sky) covered by the Aardvark Simulations. Concentric circles are lines of constant declination, 10 to $-80\degree$ in $10\degree$ increments. Radial lines are lines of constant right ascension. Opaqueness of the region indicates the overlap of catalog tiles (see text for details).} 
	\label{fig:survey area} 
\end{figure}

The ``Aardvark'' mock galaxy catalogs (R. Wechsler et al., private communication) are derived from a semi-analytical model (SAM) tied to the Bolshoi cosmological simulation \citep{Klypin2011} and designed for DES. They provide a 10313 sq. degree (one quarter of the sky) catalog to full DES depth and contains 1.36 billion galaxies which have a signal-to-noise of at least five in one or more DES observing bands. The catalog footprint is shown in Figure~\ref{fig:survey area}.

A brief description of the catalog creation is described as follows. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\tt2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\tt LGadget-2} (a version of {\tt Gadget-2}; \citealt{Springel2005}), and four different light cones are stitched together along the line of sight to create the full light cone. The DM halos are identified using the {\tt ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses and other various parameters. Galaxies are assigned to halos and subhalos using the ``ADDGals'' method of (R. Wechlar \etal, in preparation) which matches galaxy luminosities with local DM densities, not the individual DM halos. This becomes important in the velocity dispersion and subsequent dynamical mass recovery, see Section~\ref{sec:LOSVD}.

The catalog information, used in this study, is broken into two large portions. The ``truth'' files contain the characteristics of each individual galaxies, such as right ascension (RA), declination (DEC), redshift (z), observed and rest-frame magnitudes, and many others. The ``halo'' files contain information for individual halos, to which many individual galaxies may belong. This includes five estimations of dynamical mass, RA, DEC, z, three dimensional velocity dispersion, and many others.

However, the catalogs do not include information on emission or absorption lines or estimations of whether the halo is relaxed or not. We supplement the catalogs with this information and describe the method in Section~\ref{sec:oii luminosity} and others.

\subsection{Mock Observations}
\editorial{Here we need to talk about about how we set up the observations to look like what we think the HETDEX survey will see. It should be more technical than the general discussion about what the survey hopes to accomplish, that'll go in the intro. A zoomed figure showing the tiling footprints would be appropriate.}

@@

\subsection{Line of Sight Velocity Dispersion}\label{sec:LOSVD}
We first calculate the line of sight velocity (LOSV) to each galaxy, where
\begin{equation}
	LOSV = c\frac{z - z_c}{1+z_c}
\end{equation}
and $c$ is the speed of light in \kms, $z$ is the redshift of the individual galaxy, and $z_c$ is the overall cluster redshift described in the previous section.

@@

\subsection{Dynamical Mass}\label{sec:mass}
Recently, the relationship between the LOSVD and dynamical mass has been the focus of several studies \citeeg{Evrard2008, Saro2013, Sifon2013, VanderBurg2014}, and a best fitting relationship for the mass enclosed by $r_{200c}$ of the form
\begin{equation}
	M_{200c} = \frac{10^{15}}{h(z)} \bigg{(}\frac{\sigma_{1D}}{A_{1D}} \bigg{)}^{-\alpha} \Msol
\end{equation}
with $A_{1D} =$ 1040-1140 \kms\ (\citealt{Munari2013}; referred to as $\sigma_{15}$ in \citealt{Evrard2008} and other works), $\alpha = 1/3$, $h(z) = H(z)/100$, and $\sigma_{1D}$ is the LOSVD of the velocity tracers (dark matter particles, subhalos or galaxies).

@@

\section{ {\rm[\ion{O}{ii}]} Luminosity}\label{sec:oii luminosity}
HETDEX is designed to detect Lyman-$\alpha$ at redshift two. That allows us to detect [\ion{O}{ii}] ($\lambda\lambda 3727-3729\AAA$ doublet) emitters to $z\sim 0.49$. The catalog does not provide [\ion{O}{ii}] luminosities so we must assign them empirically. We use 503113 galaxies from the Sloan Digital Sky Survey (SDSS) Data Release 12 \citep{Alam2015} from $z = 0.05 - 0.2$, which are selected which have no redshift warning.
